graphsim: An R package for simulating gene expression data from graph structures of biological pathways

### 2020-06-10 paper/paper.Rmd
paper formatting
update documentation with examples of biological networks #6

Simulated datasets with pathway structure are of particular interest for
benchmarking regulatory network inference techniques and methods
being developed for genomics data with complex biological 
interactions [@Schaffter2011; @Saelens2019]. Dynamical models using
differential equations have been employed, such as by GeneNetWeaver
[@Schaffter2011], to generate simulated datasets
specifically for benchmarking gene regulatory network inference techniques.
There is renewed interest in modelling biological pathways and simulating
data for benchmarking due to the emergence of single-cell genomics
technologies and the plethora of bioinformatics techniques developed
to use this data [@Zappia2017; @Saelens2019]. Packages such as Splatter
[@Zappia2017], which uses the gamma-poisson distribution,
have been developed to model single-cell data.
SERGIO [@Dibaeinia2019] and dyngen [@Cannoodt2020] build on
this to include gene regulatory networks and multimodality
respectively. These methods have been designed based on known
or modelled reaction states and added stochasticity to determinisitic
relationships. However, it is computationally-intensive to model
these reactions at scale over iterations in benchmarking.
In some cases, it is only necessary to model the statistical
variability of RNA-Seq data to evaluate methods with these
correlation structures.

@@

#generate parameters for inhibitions for each edge in E(graph)
state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)
#plot graph structure with inhibitions (Figure 1b)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
  cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
```

@@

#generate parameters for inhibitions for each edge in E(graph)
state <- c(1, 1, -1, 1, 1, 1, 1, -1, 1)
#plot graph structure with inhibitions (Figure 1b)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
  cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
mtext(text = "(b) Inhibiting pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
box()
```

>>>>

Simulated datasets with pathway structure are of particular interest for
benchmarking regulatory network inference techniques and methods
being developed for genomics data with complex biological 
interactions [@Schaffter2011; @Saelens2019]. Dynamical models using
differential equations have been employed, such as by GeneNetWeaver
[@Schaffter2011], to generate simulated datasets
specifically for benchmarking gene regulatory network inference techniques.
There is renewed interest in modelling biological pathways and simulating
data for benchmarking due to the emergence of single-cell genomics
technologies and the plethora of bioinformatics techniques developed
to use this data [@Zappia2017; @Saelens2019]. Packages such as 'splatter'
[@Zappia2017], which uses the gamma-poisson distribution,
have been developed to model single-cell data.
SERGIO [@Dibaeinia2019] and dyngen [@Cannoodt2020] build on
this to include gene regulatory networks and multimodality
respectively. These methods have been designed based on known
or modelled reaction states and added stochasticity to determinisitic
relationships. However, it is computationally-intensive to model
these reactions at scale over iterations in benchmarking.
In some cases, it is only necessary to model the statistical
variability of RNA-Seq data to evaluate methods with these
correlation structures.

@@

#generate parameters for inhibitions for each edge in E(graph)
state <- c(1, 1, -1, 1, 1, 1, 1, -1)
#plot graph structure with inhibitions (Figure 1b)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
  cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
```

@@

#generate parameters for inhibitions for each edge in E(graph)
state <- c(1, 1, -1, 1, 1, 1, 1, -1)
#plot graph structure with inhibitions (Figure 1b)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
  cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
mtext(text = "(b) Inhibiting pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
box()
```
